---
title: " Two-Year-Olds Selectively Seek Help in the Context of Problem-Solving Analysis"
output: html_notebook
---

----------------------------EXPERIMENT 1---------------------------------------------
```{r message = FALSE, warning = FALSE}
library(readxl) #to read an excel file
library(dplyr)
library(lme4)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(tidyverse)
library(here)  
library(haven) 
library(lmerTest)  
library(splines) 
library(modelsummary) 
library(broom.mixed) 
library(simr)
library(MuMIn)
library(psych)
library(lattice)
library(sjPlot)
library(r2mlm)
library(brms)
```


# Read data

```{r}
#read data
exp1 <- read_xlsx("Experiment1Data.xlsx", range = "A1:K207")
#convert chr columns into factor
exp1 <- mutate_if(exp1, is.character, as.factor)
exp1$Subject <- as.factor(exp1$Subject)
exp1$Trial <- as.factor(exp1$Trial)
```


# Hypothesis 1

* Do kids seek more help when unfamiliar with a task compared to kids who are familiar with the task?

```{r}
# select relevant columns of data
hypo1 <- exp1 %>%
  select(Subject,Condition,Puzzle, resphelp)
str(hypo1) # There are 56 participants in total
```

## Number of data points provided by each kid

```{r}
sub <- hypo1 %>%
  group_by(Subject)%>%
  summarize(n = n())%>%
  filter(n<4)%>%
  ungroup

table(sub$n)
```



## Table of help seeking by Condition

```{r}
condition_table<-hypo1 %>%
  group_by(Condition, resphelp)%>%
  summarize(n = n())%>%
  mutate(prop = n/sum(n))%>%
  ungroup()
condition_table
```

* Based on the proportions, when unfamiliar with a task 73% of the time kids sought help compared to 27% when familiar with a task. 

## Graph of the effect of Condition

```{r}
figure<- ggplot(condition_table[condition_table$resphelp==1,], aes(x = Condition, 
                            y = prop,
                            group = 1))+
  geom_point(shape = 19,size = 3)+
  geom_line()+
  scale_x_discrete(labels = c("Familiar","Unfamiliar"))+
  ylim(0,0.8)+
  ylab("Proportion of Help Seeking")+
  theme_pubr()
figure

ggplot(condition_table, aes(x = factor(Condition), y = prop, fill = factor(resphelp)))+
  geom_col(position = "fill", width = .7)+
  scale_fill_manual(values=c("#A5A5A5","#363131"),
                    labels = c("No Help Sought","Help Sought"),
                    name = "Response")+
  scale_x_discrete(labels = c("Familiar","Unfamiliar"))+
  ylab("Proportion")+
  xlab("Condition")+
  theme_classic()

```



## Analysis: Mixed-Effects Logistic Regression

```{r}
model_help <- glmer(resphelp~Condition+(1|Subject),
                    data = hypo1,
                    control = glmerControl(optimizer = "bobyqa"),
                    family = "binomial")
summary(model_help)
confint(model_help)
```

**Result**

56 kids were randomly assigned to one of the two conditions(familiar/unfamiliar), and were asked to solve four puzzles each. For each kid and puzzle combination, we recorded whether the kid sought help (coded as 1)or did not seek help (coded as 0). To evaluate our hypothesis that kids were more likely to seek help when unfamiliar with a task, we conducted a mixed-effects logistic regression analysis with condition (familar/unfamiliar) as the fixed-effect factor. The random effect structure of the model included by-subject random intercepts. We excluded by-item random intercepts as the model with by-subject and by-item random intercepts produced a higher AIC score compared to a model with by-subject intercept only (240.1 vs 241.9). Analysis was performed using the lme4 package in R(v.4.1.3). Estimated fixed-effect parameters are given in table xx.


|  | Estimate | Std. Error | Pr(>$|z|$) | 
|-------|------|--------|-------------|  
| Intercept | -1.7887 | 0.3821 | <.0001 |  
| ConditionUnfamiliar | 1.8033 | 0.4872 | <.001 |  
 

As hypothesized, we found a significant effect of condition(familiar vs unfamiliar) in kids' help seeking behavior($\hat{\beta}=1.8033, p < .001$). This estimated parameter for the fixed-factor *condition* is in log odds-ratio. After converting log-odds ratio into odds-ratio, $e^{1.8033}=6.07$, we found that when unfamiliar with a task kids were 6 times more likely to seek for help compared to seeking help when familiar with a task.
 

# HYPOTHESIS II

* Who are kids seeking help from: baby or adult?
* Is there an effect of gender and/or sibling on kids preference of helper?

* Sibling data is in another excel file.
* Extract sibling data from the other file.

```{r}
#read the other excel file
wide_data <- read_xlsx("Data_wide.xlsx", range = "A1:Z57")
#extract sibling column
sibling <- wide_data$sib
# add sibling column to the original data set expo1
exp1$sibling <- sibling[exp1$Subject]
```

* subset kids who sought help

```{r}
# subset observations with help = 1
sub_help <- exp1%>%
  filter(resphelp == 1)%>%
  select(Subject,Gender,Puzzle,resphelpwho,sibling)
# drop unused levels
sub_help <- droplevels(sub_help)
```



```{r}
#number of times kids sought help from an adult vs a baby
with(sub_help, table(resphelpwho))
```


```{r}
# select Subject and resphelpwho from the dataset
helper <- sub_help%>%
  select(Subject,resphelpwho)
str(helper)
```

```{r}
# Testing whether there is a difference between choosing a baby or an adult as helper.
summary(glmer(resphelpwho~1+(1|Subject), data = helper, family = "binomial"))
```

* **Result**

* No significant difference in who kids are seeking help from.

### Effect of gender and/or sibling on selecting a helper


```{r}
# Create a categorical sibling column 1 = the kid has sibling, 0 = the kid does not have sibling
sub_help <- sub_help%>%
  mutate(SiblingCat = ifelse(sibling == 0,"no","yes"))
# convert SiblingCat into a factor
sub_help$SiblingCat<- as.factor(sub_help$SiblingCat)
str(sub_help)
```


```{r}
# how many datapoints with and without a sibling
with(sub_help, table(SiblingCat))
# how many male and female data points
with(sub_help, table(Gender))
# Gender by Sibling Table
with(sub_help, table(Gender, SiblingCat))
# Gender by Sibling by helper table
with(sub_help, table(SiblingCat, resphelpwho, Gender))
```


* Evaluate effect of gender and sibling on helper selection

```{r}
# effect of gender
summary(glmer(resphelpwho~Gender+(1|Subject), data = sub_help, family = "binomial"))

# effect of sibling
summary(glmer(resphelpwho~SiblingCat+(1|Subject), data = sub_help, family = "binomial"))

# effect of sibling and gender together
summary(glmer(resphelpwho~Gender+SiblingCat+(1|Subject), data = sub_help, family = "binomial"))
```

**Result**

Next, we looked at whether infants preferentially selected an adult over a baby when requesting help. We ran an intercept-only mixed effects logistic regression analysis using observations where kids sought help. The results indicated no significant difference between helper preference($\hat{\beta}=-0.01113, p = .966$). We did further analysis to test for the possible  effects that may have influenced these results. A mixed-effects logistic regression analysis was conducted to evaluate whether gender or sibling would be significant predictors of helper preference. The regression model included by-subject random intercepts. The results showed no significant effect of gender($\hat{\beta}=-0.1572, p = .774$) nor sibling($\hat{\beta}=0.5933, p = .357$).

```{r}
# summary of helper by gender and sibling
genSib <- sub_help %>%
  group_by(Gender, SiblingCat, resphelpwho)%>%
  summarize(n = n())%>%
  mutate(prop = n/sum(n))%>%
  ungroup()
genSib
```

```{r}
#graph of the helper selection by gender and sibling
ggplot(genSib, aes(x = resphelpwho, 
                   y = prop,
                   shape = SiblingCat,
                   group = SiblingCat))+
  geom_line()+
  geom_point(size = 2)+
  ylim(0,1)+
  ylab("Proportion of Help Seeking")+
  scale_x_discrete(name = "Helper", labels=c("Adult","Baby"))+
  scale_shape_discrete(name = "Sibling")+
  facet_wrap(~Gender)+
  theme_pubr()
  
```

# RQ1: In Exp 1, first time child sought help, who did they seek help from? 
### Across unfamiliar and familiar conditions (collapse them)

```{r}
# Read data
help.1 <- read.csv("Experiment1DataFT.csv", stringsAsFactors = TRUE)

# Check data
head(help.1) #shows there are columns deleted. 
tail(help.1) #shows 957 rows. This is not right.
data.frame(colnames(help.1)) #finds the index of columns. First 14 cols are relevant
summary(help.1) #shows 751 NAs for cols with actual data
# 957-751 = 206 #rows of actual data

# Clean data
help.1 <- help.1[1:206, 1:14]
str(help.1)
head(help.1) #check
tail(help.1) #check

# Fill empty cells with NAs
help.1[help.1 == "" | help.1==" "] <- NA
head(help.1)
summary(help.1)
str(help.1)

# drop unused levels of type factor in the data frame
help.1 <- droplevels(help.1)
str(help.1) #check

# Select the first observation each child sought help
first.help <- help.1 %>%
  filter(resphelp == 1)%>%
  group_by(Subject) %>%
  filter(row_number()==1)

# Table of help seeking
with(first.help, table(resphelpwho)) # A = 17, B = 19

# Get the statistical result
chisq.test(table(first.help$resphelpwho)) # not significant

## Conclusion
# We found no significant difference between who kids' sought help from(adult vs baby) 
#in their first time seeking for help (chi-square(1) = .11, p = .74)
```

----------------------------EXPERIMENT 2---------------------------------------------

```{r}
# read data
df.e2 <- read.csv("Experiment2Data.csv", stringsAsFactors = TRUE)

# convert subject, resphelp, prompted into factors
df.e2 <- mutate(df.e2, across(c("Subject", "prompted","resphelp"), factor))
```

```{r}
# clean data
# for resphelpwho and respknower, filter "" level and drop "" level
df.e2 <- df.e2 %>%
  filter(resphelpwho!="")
df.e2$resphelpwho <- droplevels(df.e2$resphelpwho)

df.e2 <- df.e2%>%
  filter(respknower!="")
df.e2$respknower<-droplevels(df.e2$respknower)
```

## RQ1: Kids seek help from an ignorant or a knowledgable helper?

```{r}
with(df.e2, table(respknower)) # I = 69, K = 107

# the analysis: intercept only mixed effects logistic regression to compare two frequencies
# ignorant = 0, knower = 1
summary(glmer(respknower~1+(1|Subject),
              data = df.e2,
              family = "binomial"))

knower <- (glmer(respknower~1+(1|Subject),
              data = df.e2,
              family = "binomial"))

confint(knower)

```

* Results from mixed-effects logistic regression showed that when unfamiliar with a task kids are 1.55 times (exp(0.4387)) more likely to seek help from someone who is knowledgeable about the task compared to someone who is not(z = 2.842, p = .005). 

### Does the type of helper (child vs adult) have an effect on whether they seek from a knowledgeable helper or an ignorant helper?

```{r}
with(df.e2, table(respknower, resphelpwho)) # it does not look so from the table
```

```{r}
summary(glmer(respknower~resphelpwho+(1|Subject),
              data = df.e2,
              family = "binomial"))

helper <- (glmer(respknower~resphelpwho+(1|Subject),
              data = df.e2,
              family = "binomial"))

confint(helper)
```

```{r}
# RQ 2: Same as RQ 1 but for experiment 2 (adult vs peer)

# read data
help.2 <- read.csv("Experiment2DataFT.csv", stringsAsFactors = TRUE)

# Check data
head(help.2) # has extra columns
data.frame(colnames(help.2)) # get index of col names
tail(help.2) # goes all the way to 965 rows
summary(help.2)
#965-784 #There are 181 rows of data 

# Clean data
help.2 <- help.2[1:181, 1:19]
summary(help.2) # check

# Replace empty cells with NAs
help.2[help.2 == "" | help.2==" "] <- NA
summary(help.2) #check

# drop unused levels of factors
help.2 <- droplevels(help.2)
summary(help.2) #check

# Select the first observation each child sought help
help.exp2 <- help.2 %>%
  filter(resphelp == 1)%>%
  group_by(Subject) %>%
  filter(row_number()==1)

# Table of help seeking
with(help.exp2, table(resphelpwho)) #A = 22, S = 26

# Get the statistical result
chisq.test(table(help.exp2$resphelpwho))

## Conclusion
# We found no significant difference between who kids' sought help from(adult vs peer) 
#in their first time seeking for help (chi-square(1) = .33, p = .56)
```

```{r}
# Experiment 2
# Simple Effect of Gender 

## Freq distribution of help seeking by gender(whether they sought help or not)
with(help.2, table(Gender, resphelp))
# There shouldn't be 0's as kids are forced to seek help. Filter resphelp == 1
help.help <- help.2 %>%
  filter(resphelp == 1)

with(help.help, table(Gender, resphelp)) # check there are no 0s.

## Freq distribution of who kids seek help from by gender
with(help.help, table(Gender, resphelpwho)) # does not look different

## Run mixed effects logistic regression to verify 
m.gender <- glmer(resphelpwho~Gender+(1|Subject), data = help.help, family = "binomial")
summary(m.gender)
# No significant effect of gender on who kids seek help from.

```

```{r}
## EFFECT OF SIBLING ON HELP SEEKING EXPERIMENT 2 (A vs S)
with(help.2, table(sib, resphelpwho))
# Frequency distribution shows more help seeking from an adult when they no sibling
summary(glmer(resphelpwho~sib+(1|Subject),
              data = help.2,
              family = "binomial")) # not significant
```

----------------------------EXPERIMENT 3---------------------------------------------
```{r}
exp3 <- read.csv("Experiment3Data.csv") #load readxl package to use this str(exp3)
str(exp3)

anyNA(exp3)

nas <- !complete.cases(exp3)#returns a logical vector. TRUE if a row has NA, FALSE otherwi which(nas)
which(nas)

exp3 <- exp3[-c(115,131),]

str(exp3)

anyNA(exp3)

exp3 <- exp3%>%
  select(Subject, TempOrdWho, RespHelpWho)
str(exp3)

exp3 <- exp3 %>%
    mutate(Response = ifelse(RespHelpWho == "A", 1, 0),
           Condition = ifelse(TempOrdWho == "B", 0, 1))

with(exp3, table(Response, Condition))

m1 <-glmer(Response~Condition + (1|Subject),
           data = exp3, family = "binomial")

summary(m1)

confint(m1)
```

```{r}
# For Experiment 3: Look at help seeking for Adult vs non-adult
## The intercept only model
### RQ: Is it the case that kids seek more help from an adult vs non-adult?
with(help.3.sub, table(HelpWho))
# it actually looks like kids seek more help from a nonadult (0 category)
summary(glmer(HelpWho~1+(1|Subject),
              data = help.3.sub,
              family = "binomial"))

exp3.pooled <- (glmer(HelpWho~1+(1|Subject),
              data = help.3.sub,
              family = "binomial"))


help.3.sub$HelpWho <- as.factor(help.3.sub$HelpWho)
str(help.3.sub)
summary(exp3.pooled)
confint(exp3.pooled)

## the difference is statistically not significant (b = -.24, z = -1.28, p = .2)
```


```{r}
## EFFECT OF SIBLING ON HELP SEEKING (A VS B) AND (A VS S) SEPARATELY. EXPERIMENT 3
with(help.3.sub, table(TempOrdWho,RespHelpWho,Sibling))

# Subset data with TempOrdWho == B (This is to compare A vs B give the choice A vs B)
help.3.AvsB <- help.3.sub %>%
  filter(TempOrdWho == "B")
# drop unusued levels
help.3.AvsB <- droplevels(help.3.AvsB)
str(help.3.AvsB)
# Regress RespHelpWho on Sibling
summary(glmer(HelpWho~Sibling+(1|Subject),
              data = help.3.AvsB,
              family = "binomial")) # not significant

# Subset data with TempOrdWho == S (This is to compare A vs S given the choice A vs S)
help.3.AvsS <- help.3.sub %>%
  filter(TempOrdWho == "S")
#drop unused levels
help.3.AvsS <- droplevels(help.3.AvsS)
str(help.3.AvsS)
# Regress RespHelpWho on Sibling
summary(glmer(HelpWho~Sibling+(1|Subject),
              data = help.3.AvsS,
              family = "binomial"))
## Conclusion: Given the option to choose help from an adult versus a baby
# or adult from a peer, kids who did not have a sibling sought help more from an adult, 
# while kids who did have a sibling sought more help from a baby or a peer. 
# However, these findings were statistically not significant.

# EXPERIMENT 3 EFFECT OF GENDER 
# Regress RespHelpWho on GENDER A VS B
summary(glmer(HelpWho~Gender+(1|Subject),
              data = help.3.AvsB,
              family = "binomial")) # not significant

# Regress RespHelpWho on GENDER A vs S
summary(glmer(HelpWho~Gender+(1|Subject),
              data = help.3.AvsS,
              family = "binomial")) # not significant

# CONCLUSION: no effect of gender
```

